---
title: "Single Nuclei vs Single cells (non-depleted) -- Normal tissue, 2 patients"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
params:
  datapath: 'C:/Users/renseb01/Documents/scRNA/scRNA/data'
  outputpath: '../..' 
---

```{r, timehook,echo=F}
knitr::knit_hooks$set(timehook = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("\n\n\nTime for this code chunk to run:",signif(res,3),attr(res,'units'))
    }
  }
}))
```

```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = params$datapath)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE, cache.lazy=FALSE)
```

```{r, requirements}
#requirements
library(ggalluvial)
library(dplyr)
library(ggplot2)
library(patchwork)
library(Seurat)

library(readxl)
```

# Loading data  
```{r loadRDS,message  = T,echo=T}  
adeno.combined.sct =  readRDS('adeno.combined.sct.rds')

#summarise your data
data_summarised = adeno.combined.sct@meta.data %>% 
  group_by(predicted.ann_level_1,predicted.ann_level_2,predicted.ann_level_3,predicted.ann_finest_level,method) %>%
  summarise(n = length(predicted.ann_level_2),.groups='keep') %>% 
  filter(n > 50)

data_summarised$method = ifelse(data_summarised$method=='SN','SingleNuclei','SingleCell')
```

# Cell type annotation 
```{r celltypes,fig.width = 13, fig.height=7,echo=F}
# Stacked barplot with multiple groups
high_level_colors = c(RColorBrewer::brewer.pal(n = 5, name = "Set1"))
bargplots = list()
bargplots[[1]] = ggplot(data=data_summarised, aes(x=method, y=n, fill=predicted.ann_level_1)) +
  geom_bar(stat="identity") + 
  scale_fill_manual('Cell Types',values=high_level_colors)+ 
  ggtitle('Cell Types -- coarse annotation') + ylab("Number of Cells")
  
#
celltypes = c('Endothelial','Epithelial','Immune','Stroma')

for(i in seq_along((celltypes)))
  {
  data = data_summarised[data_summarised[,1]==celltypes[i],]
  rgb.palette <- colorRampPalette(c('black',high_level_colors[i], "white"),space = "rgb")
  colors = rgb.palette(length(unique(data$predicted.ann_finest_level))-1)
  colors[length(colors)+1] = 'orange'

  p = ggplot(data=data, aes(x=method, y=n, fill=predicted.ann_finest_level)) +
    geom_bar(stat="identity") +
    scale_fill_manual('Cell Types',values=colors)+ 
    ggtitle(celltypes[i])+ ylab("")

  bargplots[[i+1]] = p
}

bargplots[[1]]+(bargplots[[2]]/bargplots[[3]])+(bargplots[[4]]/bargplots[[5]])

#pdf('results/cell_type_annotation.pdf',width = 13,height = 7)
#bargplots[[1]]+(bargplots[[2]]/bargplots[[3]])+(bargplots[[4]]/bargplots[[5]])
#dev.off()

```


# Proof-Of-Concept for DEGs
 * Not shown. A POC to show how `Seurat` looks for differentially expressed genes.  
```{r deg POC, echo=F, message=FALSE, include=FALSE}
# Take all cells in cluster 2, and find markers that separate cells in the 'g1' group (metadata
# variable 'group')
markers <- FindMarkers(pbmc_small, slot = 'data',assays = 'RNA', ident.1 = "g1", group.by = 'groups', subset.ident = "2")

head(x = markers,1)

g1 = pbmc_small@assays$RNA@data[,pbmc_small@meta.data$RNA_snn_res.1 == 2 & pbmc_small@meta.data$groups=='g1']

g2 = pbmc_small@assays$RNA@data[,pbmc_small@meta.data$RNA_snn_res.1 ==2 & pbmc_small@meta.data$groups=='g2']

g1_1 = g1[rownames(g1) == "GSTP1", ]
g2_1 = g2[rownames(g2) == "GSTP1", ]
pct.1 = length(g1_1[g1_1!=0])/length(g1_1)
pct.2 = length(g2_1[g2_1!=0])/length(g2_1)
FC = log(mean(expm1(g1_1))+1,2)-log(mean(expm1(g2_1))+1,2)
pval = wilcox.test(g1_1,g2_1)$p.value

# find markers for every cluster compared to all remaining cells, report only the positive ones
adeno.combined.sct <- SetIdent(adeno.combined.sct, value = 'predicted.ann_finest_level')

markers <- FindMarkers(adeno.combined.sct, ident.1 = "P4",ident.2 = "P6", group.by = 'patient', subset.ident = 'NK cells')
markers1 <- FindMarkers(adeno.combined.sct, ident.1 = "SN",ident.2 = "SC", group.by = 'method', subset.ident = 'AT1')
markers2 <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "SN",ident.2 = "SC", group.by = 'method', subset.ident = 'AT2',assay = 'integrated')
markers3 <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "SN",ident.2 = "SC", group.by = 'method', assay = 'integrated')

###PCT.1 AND PCT.2 ARE  TRUE, i now  get the Fold Change.
PGC_rawdata = adeno.combined.sct@assays$integrated@data[rownames(adeno.combined.sct@assays$integrated@data)=='PGC',]
PGC_rawdata_AT2_SN = PGC_rawdata[adeno.combined.sct@meta.data$predicted.ann_finest_level=='AT2' & adeno.combined.sct@meta.data$method=='SN']
PGC_rawdata_AT2_SC = PGC_rawdata[adeno.combined.sct@meta.data$predicted.ann_finest_level=='AT2' & adeno.combined.sct@meta.data$method=='SC']

#FindMarkers statistics
markers2[rownames(markers2)=='PGC',]

#hand calculation
length(PGC_rawdata_AT2_SC[PGC_rawdata_AT2_SC>0])/length(PGC_rawdata_AT2_SC) 
length(PGC_rawdata_AT2_SN[PGC_rawdata_AT2_SN>0])/length(PGC_rawdata_AT2_SN) 

#scale.data
log(mean((PGC_rawdata_AT2_SN))+1,2)-log(mean((PGC_rawdata_AT2_SC))+1,2)

#if data
log(mean(expm1(PGC_rawdata_AT2_SN))+1,2)-log(mean(expm1(PGC_rawdata_AT2_SC))+1,2)

```




# DEGs  
We can notice five things:   
1. In the 1st panel, we can see that the expression in centered around zero, which means that the __gene * cell__ matrix is very sparse.  
2. In the 1st panel, AT1 and AT2 marker genes (used to identify them as AT1/AT2 cells, as defined by **HLCA**) have overall low expression.  
3. In the second panel, we can notice that, in AT1 cells specifically, the expression of marker genes (as defined by *HLCA*) is very high. While this is true in Single Nuclei, it is even higher in Single Cells.  
4. In the second panel, we also notice that, in AT1 cells, the overall expression of genes is higher than in all cells, both in Single Cells and Single Nuclei. While this is true in Single Nuclei, it is even higher in Single Cells.  
5. Patterns *3* and *4* are the same for AT2 cells.  
```{r degs,fig.width = 11, fig.height=5, echo=F}
markers = read.csv('marker_genes_April2023.txt',header = F,row.names = 1)

output = data.frame()

for(i in c(3,12)){
  meta = adeno.combined.sct@meta.data
  alldata = adeno.combined.sct@assays$integrated@scale.data
  
  celltypeX_SN = alldata[,meta$method == 'SN' & meta$predicted.ann_finest_level == rownames(markers)[i]]
  celltypeX_SC = alldata[,meta$method == 'SC' & meta$predicted.ann_finest_level == rownames(markers)[i]]
  
  celltypeX_SN_markers = celltypeX_SN[rownames(celltypeX_SN) %in% markers[i,],]
  celltypeX_SN_nonmarkers = celltypeX_SN[!(rownames(celltypeX_SN) %in% markers[i,]),]
  
  celltypeX_SC_markers = celltypeX_SC[rownames(celltypeX_SC) %in% markers[i,],]
  celltypeX_SC_nonmarkers = celltypeX_SC[!(rownames(celltypeX_SC) %in% markers[i,]),]
  
  output_temp = data.frame(expression = c(as.vector(celltypeX_SN_markers),
                   as.vector(celltypeX_SN_nonmarkers),
                   as.vector(celltypeX_SC_markers),
                   as.vector(celltypeX_SC_nonmarkers)))
  
  output_temp$method = c(rep('Single Nuclei',length(as.vector(celltypeX_SN_markers))),
                          rep('Single Nuclei',length(as.vector(celltypeX_SN_nonmarkers))),
                          rep('Single Cell',length(as.vector(celltypeX_SC_markers))),
                          rep('Single Cell',length(as.vector(celltypeX_SC_nonmarkers)))
                          ) 
  
  output_temp$categories = c(rep('marker genes',length(as.vector(celltypeX_SN_markers))),
                          rep('non-marker genes',length(as.vector(celltypeX_SN_nonmarkers))),
                          rep('marker genes',length(as.vector(celltypeX_SC_markers))),
                          rep('non-marker genes',length(as.vector(celltypeX_SC_nonmarkers)))
                          )
  
  output_temp$predicted.ann_finest_level = paste0(rownames(markers)[i], ' cells')
  
  output_temp = output_temp[output_temp$expression>0,]
  
  output = rbind(output,output_temp)
  
  if(i ==12) {
  SN = alldata[,meta$method == 'SN']
  SC = alldata[,meta$method == 'SC']
  
  celltypeX_SN_markers_at2 = SN[rownames(SN) %in% markers[3,],]
  celltypeX_SN_markers_at1 = SN[rownames(SN) %in% markers[12,],]

  celltypeX_SC_markers_at1 = SC[rownames(SC) %in% markers[12,],]
  celltypeX_SC_markers_at2 = SC[rownames(SC) %in% markers[3,],]

  output_temp = data.frame(expression = c(as.vector(celltypeX_SN_markers_at2),
                                          as.vector(celltypeX_SN_markers_at1),
                                          as.vector(SN),
                                          as.vector(celltypeX_SC_markers_at2),
                                          as.vector(celltypeX_SC_markers_at1),
                                          as.vector(SC)),
                                          
                           
                             method=c(rep('Single Nuclei',length(as.vector(celltypeX_SN_markers_at2))),
                                      rep('Single Nuclei',length(as.vector(celltypeX_SN_markers_at1))),
                                      rep('Single Nuclei',length(as.vector(SN))),
                                      rep('Single Cell',length(as.vector(celltypeX_SC_markers_at2))),
                                      rep('Single Cell',length(as.vector(celltypeX_SC_markers_at1))),
                                      rep('Single Cell',length(as.vector(SC)))),
                           
                           
                             categories=c(rep('AT2 marker genes',length(as.vector(celltypeX_SN_markers_at2))),
                                      rep('AT1 marker genes',length(as.vector(celltypeX_SN_markers_at1))),
                                      rep('All genes',length(as.vector(SN))),
                                      rep('AT2 marker genes',length(as.vector(celltypeX_SC_markers_at2))),
                                      rep('AT1 marker genes',length(as.vector(celltypeX_SC_markers_at1))),
                                      rep('All genes',length(as.vector(SC)))),
                           
                             predicted.ann_finest_level = 'all cells')
    
    output_temp = output_temp[sample(1:40000000,5000000),]
    output = rbind(output,output_temp)
  }
  
}

###plot
violon_plots = ggplot(output, aes(x=categories, y=expression, fill=method)) +
  geom_violin(trim=FALSE,position="dodge") +
  ylab("Normalised gene expression") +
  facet_wrap(~predicted.ann_finest_level,nrow = 1, ncol = 5, scales = "free_x")
###

violon_plots


pdf('../results/violon_plots.pdf',width = 13,height = 7)
violon_plots
dev.off()

```

# DEGs -- part2
We can notice four things:  
1. In the first (top left) panel, there are hundreds of genes differentially expressed (DEGs) between Single Nuclei and Single Cells.  
2. In the second panel (top right), we can see that, overall, genes are more expressed in Single Cells (negative values), but these mean differences are hard to interpret because they are *SCT* normalized data from a sparse matrix.  
3. In the third panel (bottom left), overall DEGS are expressed in ~13% of the cells for Single cells vs ~25% for Single Nuclei.  
4. However, as expected, AT1 & AT2 marker genes are expressed in a very large fraction of AT1 and AT2 cells. All other DEGs are expressed in a slightly larger fractions of AT1 and AT2 specific cells than all cells combined.   
   
```{r degs part2,eval=T,fig.width = 13, fig.height=7,echo=F,message =F}
adeno.combined.sct <- SetIdent(adeno.combined.sct, value = 'predicted.ann_finest_level')

deg_at1 <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "SN",ident.2 = "SC", group.by = 'method', subset.ident = 'AT1',assay = 'integrated')
deg_at1 = deg_at1[deg_at1$p_val_adj<0.05,]
deg_at1$categories = 'non-marker genes'
deg_at1$categories[rownames(deg_at1) %in% unlist(markers[c(12),])] = 'marker genes'

deg_at2 <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "SN",ident.2 = "SC", group.by = 'method', subset.ident = 'AT2',assay = 'integrated')
deg_at2 = deg_at2[deg_at2$p_val_adj<0.05,]
deg_at2$categories = 'non-marker genes'
deg_at2$categories[rownames(deg_at2) %in% unlist(markers[c(3),])] = 'marker genes'

deg_all <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "SN",ident.2 = "SC", group.by = 'method', assay = 'integrated')
deg_all = deg_all[deg_all$p_val_adj<0.05,]
deg_all$categories = 'all'

deg = data.frame(rbind(deg_all,deg_at1,deg_at2),predicted.ann_finest_level = c(rep('All cells',nrow(deg_all)),rep('AT2 cells',nrow(deg_at1)),rep('AT1 cells',nrow(deg_at2))))

#summarise your data
data_summarised = deg %>% 
  group_by(categories,predicted.ann_finest_level) %>%
  summarise(n = length(predicted.ann_finest_level),
            mean_SC = mean(pct.2),
            mean_SN = mean(pct.1),
            avg_diff = mean(avg_diff),.groups='keep')

#add a single pct column
nrows = nrow(data_summarised)
data_summarised = rbind(data_summarised,data_summarised)
data_summarised$mean_SC = c(data_summarised$mean_SC[1:nrows],data_summarised$mean_SN[1:nrows])
data_summarised$mean_SN = c(rep('Single Cells',nrows),rep('Single Nuclei',nrows))
data_summarised$n[(nrows+1):nrow(data_summarised)] = 0
data_summarised$avg_diff[(nrows+1):nrow(data_summarised)] = 0
colnames(data_summarised)[4] = 'pct' 
colnames(data_summarised)[5] = 'method'
data_summarised$categories2 = data_summarised$categories

data_summarised$categories2[c(1,6)] = 'all DEGs (1567 genes)' 
data_summarised$categories2[c(2,7)] = 'markers (7 DEGs)' 
data_summarised$categories2[c(3,8)] = 'markers (9 DEGs)' 
data_summarised$categories2[c(4,9)] = 'non-markers DEGs\n(693 genes)'  
data_summarised$categories2[c(5,10)] = 'non-markers DEGs\n(1142 genes)'  
 
p_nb_deg = ggplot(data=data_summarised, aes(x=predicted.ann_finest_level, y=n)) +
    geom_bar(stat="identity") +
    ggtitle('Number of Differentially Expressed Genes (DEGs)\n(Single Nuclei vs Single Cells)') +
  ylab("Number of Genes") + 
  xlab("")

p_avg_diff =ggplot(deg, aes(x=predicted.ann_finest_level, y=avg_diff,fill = categories)) + 
  geom_boxplot(position=position_dodge(1)) +
  ggtitle('Mean difference in normalized expression\n(Single Nuclei - Single Cells)') + 
  ylab("Mean difference") + 
  xlab("")

p_pct = ggplot(data=data_summarised, aes(x=categories2, y=pct, fill= method)) +
    geom_bar(stat="identity", position=position_dodge()) +
    ggtitle('Mean fraction of cells expressing a DEG') + 
  facet_wrap(~predicted.ann_finest_level,nrow = 1, ncol = 3, scales = "free_x") +
  ylab("Percentage") +
  xlab("Cell Types")

#plot
(p_nb_deg + p_avg_diff) / p_pct 


pdf('../results/DEG.pdf',width = 13,height = 7)
(p_nb_deg + p_avg_diff) / p_pct 
dev.off()


```


