---
title: "scRNA lung adenocarcinoma"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
params:  
  datapath: 'C:/Users/renseb01/Documents/scRNA/scRNA_lung_adeno/data/' 
  nFeature_min: 250  
  nCount_min: 500  
  nCount_max: 50000  
  percent_MT_max: 2  
  
---

## Introduction  
* This is a nucleus - whole cell comparison. Here, we have a *eight* matrices (generated with `cellranger`). For each sample we have 4 conditions:   
  * Tumor / Normal tissue * immune cells / depleted.  
* Below is a short Proof-Of-Concept of analyses to be done using the `seurat` R package. It mainly follows the [tutorials explained here](https://satijalab.org/seurat/articles/get_started.html).
* Currently, the goal of the project is to: 
  *  compare *NEVER* vs *EVER* smokers
  *  compare *tumor* vs *normal* tissue
  *  compare *total tissue* vs *immune cell depleted* samples  

## R markdown setup 
  * setup, requirements, default knitr options, rootdir, timehook  
```{r, timehook,echo=F}
knitr::knit_hooks$set(timehook = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("\n\n\nTime for this code chunk to run:",signif(res,3),attr(res,'units'))
    }
  }
}))
```

```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = params$datapath)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE, cache.lazy=FALSE,timehook=TRUE)
```



```{r, requirements}
library(dplyr) #data handling
library(Seurat) #scRNA
library(patchwork) #data viz
library(sctransform) #data normalization
library(ggplot2) #data viz
library(DT) #data viz
library(scAnnotatR) #cell type annotation (WARNING: this is currently installed via devtools::install_git('griss/scAnnotatr@develop'))
```


## Loading datasets and merging  
  * `barcodes.tsv.gz`, `features.tsv.gz` and `matrix.mtx.gz` files come directly from  `cell ranger`. `aggregation.csv` file specify the samples and experimental conditions.
  
```{r initial_load}
# Load the cellranger datasets
#Multiple samples stored in filtered_feature_bc_matrix
whole_cells = 'C:/Users/renseb01/Documents/scRNA/scRNA_lung_adeno/data/aggr3018_3019_3027_3028/filtered_feature_bc_matrix'
nuclei = 'C:/Users/renseb01/Documents/scRNA/scRNA_lung_adeno/data/results_aggr-3018-3019-3027-3028_nuclei/filtered_feature_bc_matrix' 
datapath = c(whole_cells,nuclei)

#create seurat objects
seurat.objects = list()
for(i in 1:length(datapath)) {
#for(i in 1:4) {
  cts <- Read10X(data.dir = datapath[i])
  seurat.objects[[i]] = CreateSeuratObject(counts = cts)

aggr = read.csv(file.path(datapath[i],'aggregation.csv'))
 
 if(i ==1){
  #sample specific IDs are here:WHOLE CELLS
  aggr$method = rep(c('all','depleted'),4)
  aggr$patient = c(rep(c('P4'),4),rep(c('P6'),4))
  aggr$samplename = unlist(sapply( strsplit(aggr$sample_id,split="_"), "[", 1))
  aggr$sampletag = 1:8
  aggr$type = c('normal','normal','tumor','tumor','normal','normal','tumor','tumor')
  aggr$organelle = rep('wholecell',4)
  aggr$organelleXtype = paste0(aggr$organelle,'_',aggr$type)
  }
 
  #nuclei metadata 
  if(i ==2){
  aggr$method = rep('all',4)
  aggr$patient = c(rep(c('P4'),2),rep(c('P6'),2))
  aggr$samplename = unlist(sapply( strsplit(aggr$sample_id,split="_"), "[", 1))
  aggr$sampletag = 1:4
  aggr$type = c('normal','tumor','normal','tumor')
  aggr$organelle = rep('nuclei',4)
  aggr$organelleXtype = paste0(aggr$organelle,'_',aggr$type)
  }

  #fill the metadata
  seurat.objects[[i]][["sampletag"]] = unlist(sapply(strsplit(rownames(seurat.objects[[i]]@meta.data),split="-"), "[", 2))
  merged_meta = merge(seurat.objects[[i]]@meta.data,aggr[,-2],by = "sampletag")
  seurat.objects[[i]]@meta.data[,1:ncol(merged_meta)] = merged_meta
  colnames(seurat.objects[[i]]@meta.data) = colnames(merged_meta)
  head(seurat.objects[[i]]@meta.data)
}


#merge datasets
adenocarcino <- merge(seurat.objects[[1]],
                       y = seurat.objects[-1],
                       add.cell.ids = c('c','n'),
                       project='irsc')

adenocarcino
```

# Quality control  
```{r qc,message  = T, fig.width = 11, fig.height=8,echo=F}
#filters
message(paste0('Applying filters here: \n nFeature_RNA_min: ',params$nFeature_min,
               '\n nCount_min: ',params$nCount_min,
               '\n nCount_max: ',params$nCount_max,
               '\n percent.mt: ',params$percent_MT_max))

#add MT fractions to the metadata
adenocarcino[["percent.mt"]] <- PercentageFeatureSet(adenocarcino, pattern = "^MT-")

# Visualize QC metrics as a violin plot
plot_violon = VlnPlot(adenocarcino, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, log = T,group.by = 'organelle')
plot_violon


#subset based on QC metrics
adenocarcino_subset <- subset(adenocarcino, subset = 
                                nFeature_RNA > params$nFeature_min & 
                                nCount_RNA < params$nCount_max &
                                nCount_RNA > params$nCount_min &
                                percent.mt < params$percent_MT_max &
                                method == 'all')
```

## Clean Up: Before & After  
```{r before after,message=TRUE, echo=FALSE}
message('---seurat object before clean-up---')
adenocarcino

message('---seurat object after clean-up---')
adenocarcino_subset
```


# SCTranform  
  * This step is performed instead of the scaling, normalization & feature selection steps.  
  * It removes unwanted effects (e.g. MT effect) from UMI data by modelling them, and keeping the residuals.
  * It's based on a *Variance Stabilization Transformation* (VST). So like a log transformation, but also dealing with the *expression variance --- expression mean* correlation (i.e. heteroscedasticity).  
```{r SCTransform}
#subset based on QC metrics
adenocarcino_subset <- SCTransform(adenocarcino_subset, method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = F)
```


# Dimensionality reduction (PCA)   
  * This step is performed in order to reduce the dimensionality of the dataset from several thousands genes to a relatively small number of (orthogonal) Principal Components (*PCs*). PCs are ranked in decreasing order based on the percentage of variance they explain (PVE). Ideally PVE decreases logarithmically such that the first few PCS explain most of the variance. 
  * Here, we need to assess visually how many PCs we want to keep in order to cluster samples in the next step.  
```{r PCA,fig.width = 11}
adenocarcino_subset <- RunPCA(adenocarcino_subset,npcs = 50)

#dimensionality reduction (PCA)
plot_pca_patient = DimPlot(adenocarcino_subset, reduction = "pca", group.by = 'patient')
plot_pca_type = DimPlot(adenocarcino_subset, reduction = "pca", group.by = 'type')
plot_pca_method = DimPlot(adenocarcino_subset, reduction = "pca", group.by = 'organelle')
plot_pca_patient + plot_pca_type + plot_pca_method

#heatmaps of the first 12 PCs...
plot_n_heatmap = DimHeatmap(adenocarcino_subset, dims = 1:12, cells = 500, balanced = TRUE, fast = F)
plot_n_heatmap

#Percentage of variance explained by the first 50 PCs.
scaled_data <- Seurat::GetAssayData(adenocarcino_subset, assay = "SCT", slot = "scale.data")
total_variance <- sum(matrixStats::rowVars(scaled_data))
eig_values = (adenocarcino_subset@reductions$pca@stdev)^2  ## EigenValues
pca_stats = data.frame(PC = 1:length(eig_values),var = sqrt(eig_values), pve = signif(eig_values / total_variance *100,2), pve = signif(eig_values / total_variance *100,2))

ggplot(data=pca_stats, aes(x=PC, y=var)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=pve), vjust=1.6, color="darkblue",
            position = position_dodge(0.9), size=3.5)+
  labs(title=paste0("PCA elbow plot & (PVE in blue, first ",nrow(pca_stats)," PCs explain ",signif(sum(pca_stats$pve),3),"% of variance)"),
         y="Variance", x = "Principal Components")
```


# Graph construction, optimal Nb of clusters, UMAP  
  * This step embeds cells in a K-nearest neighbor (KNN) graph, with edges drawn between cells with similar feature expression patterns.  
  * Cells are then partitioned into highly interconnected ‘clusters’ (based on Louvain algorithm which optimises the modularity score for each ‘clusters’). *Note that the optimal number of clusters is dependent of the number of PCs chosen AND the granularity (resolution). I currently keep 20 PCs, but try 3 different cluster resolution values*.  
  *  Finally, we plot similar cells together in 2-dimensional space (UMAP) based on a set number of PCs.  

```{r UMAP, fig.width = 11, fig.height=8}
adenocarcino_subset <- FindNeighbors(adenocarcino_subset, reduction = "pca", dims = 1:20)
adenocarcino_subset <- FindClusters(adenocarcino_subset, resolution = 0.5)
adenocarcino_subset <- FindClusters(adenocarcino_subset, resolution = 0.8)
adenocarcino_subset <- FindClusters(adenocarcino_subset, resolution = 1)
adenocarcino_subset <- FindClusters(adenocarcino_subset, resolution = 1.5) 

adenocarcino_subset <- RunUMAP(adenocarcino_subset, reduction = "pca", dims = 1:20, verbose = F)
plot_umap = DimPlot(adenocarcino_subset, reduction = "umap",label = T,group.by =  'SCT_snn_res.0.5')
plot_umap

#lets use the 0.5 granularity for the rest of the analyses...
adenocarcino_subset@active.ident = adenocarcino_subset@meta.data$SCT_snn_res.0.5
names(adenocarcino_subset@active.ident) =  rownames(adenocarcino_subset@meta.data)

```


# Plot UMAPs according to specific effects.
  * **Patient**, **type**, **method**, **method X type** effects  
```{r specific effects, fig.width = 11, fig.height=8}

p1 <- DimPlot(adenocarcino_subset, reduction = "umap", group.by = "patient",cols=paste0(DiscretePalette(16)[1:2],c('BB','22')),size =0.1)
p2 <- DimPlot(adenocarcino_subset, reduction = "umap", group.by = "type",cols = paste0(DiscretePalette(16)[7:8],c('BB','22')),size = 0.1)
p3 <- DimPlot(adenocarcino_subset, reduction = "umap", group.by = "organelle",cols=paste0(DiscretePalette(16)[c(14,16)],c('BB','22')),size = 0.1)
p4 <- DimPlot(adenocarcino_subset, reduction = "umap", group.by = "organelleXtype",cols=paste0(DiscretePalette(16)[c(9,10,13,15)],c('BB','22','22','22')),size =0.1)
p1 + p2 + p3 + p4 
```

# Assigning cell type identity to clusters  
  * Read this [paper](https://bmcbioinformatics.biomedcentral.com/articles/10.1186/s12859-022-04574-5) for methodology.
  * This is really a Work-In-Progress, as we only currently use about 40 very general cell types. 
    * Immune-related clusters are nicely annotated with immune cells (Bcells, Tcells, etc).  
    * Note also the large fraction of unknown cell types.  
```{r assessing cell type,message=F,echo=T,eval=T,fig.width = 11,fig.height=8}
###
default_models <- scAnnotatR::load_models("default")
adenocarcino_subset <- classify_cells(classify_obj = adenocarcino_subset, 
                             assay = 'SCT', slot = 'counts',
                             cell_types = 'all', 
                             path_to_models = 'default')

#summarise, removing unknown assignments...
ct_summarised <- adenocarcino_subset@meta.data %>%
  group_by(most_probable_cell_type,SCT_snn_res.0.5,.drop=F) %>% 
  summarise(cluster = length(SCT_snn_res.1)) %>%
  filter(most_probable_cell_type!='unknown')

plot_cpct = ggplot(data=ct_summarised, aes(y=cluster, x=most_probable_cell_type,fill = SCT_snn_res.0.5)) + 
  geom_bar(stat="identity") + 
  labs(title= 'nb cells per cell type',y = "Nb of cells", x = "cell type") + 
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  guides(fill = guide_legend(nrow = 10))

plot_cpc = ggplot(data=ct_summarised, aes(y=cluster, x=SCT_snn_res.0.5, fill = most_probable_cell_type)) + 
    geom_bar(stat="identity") + 
    labs(title= 'nb cells per cluster',y = "Nb of cells", x = "cell clusters") + 
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    guides(fill = guide_legend(nrow = 10))

plot_cpct + plot_cpc + plot_layout(nrow=2)


```

```{r DT_most_prob_celltype,message = T, echo = F}
temp = rle(sort(adenocarcino_subset@meta.data$most_probable_cell_type))
mpct = data.frame(type = temp$values, count = temp$lengths)
datatable(mpct)
```


# Quantifying the proportions of cells  
  * A simple visualisation to see how cell populations differ between groups.  
```{r proportions,fig.width = 11, fig.height=8}
###proportions of cells
summarised <- adenocarcino_subset@meta.data %>% group_by(patient, type, organelle,organelleXtype,SCT_snn_res.0.5,.drop=F) %>% summarise(cluster = length(SCT_snn_res.0.5))
summarised$proportions = summarised$cluster /sum(summarised$cluster) *100

plot_proportions1 <- ggplot(data=summarised, aes(y=proportions, x=SCT_snn_res.0.5,fill = patient)) + 
  geom_bar(stat="identity",position=position_dodge()) + 
  scale_fill_manual(values=DiscretePalette(16)[1:2]) + 
  labs(title= 'cell proportions per patient',y = "% of cells in cluster", x = "clusters")

plot_proportions2 <- ggplot(data=summarised, aes(y=proportions, x=SCT_snn_res.0.5,fill = type)) + 
  geom_bar(stat="identity",position=position_dodge()) + 
  scale_fill_manual(values=DiscretePalette(16)[7:8]) + 
  labs(title= 'cell proportions per type',y = "% of cells in cluster", x = "clusters")

plot_proportions3 <- ggplot(data=summarised, aes(y=proportions, x=SCT_snn_res.0.5,fill = organelle)) + 
  geom_bar(stat="identity",position=position_dodge()) +
  scale_fill_manual(values=DiscretePalette(16)[c(14,16)]) + 
  labs(title= 'cell proportions per method',y = "% of cells in cluster", x = "clusters")

plot_proportions4 <- ggplot(data=summarised, aes(y=proportions, x=SCT_snn_res.0.5,fill = organelleXtype)) + 
  geom_bar(stat="identity",position=position_dodge()) +
  scale_fill_manual(values=DiscretePalette(16)[c(9,10,13,15)]) + 
  labs(title= 'cell proportions per method',y = "% of cells in cluster", x = "clusters")

plot_proportions1 + plot_proportions2 + plot_proportions3 + plot_proportions4
```

# Top DEGs  
  * This compares each cluster seperately against all others and looks for diferentially expression genes. It then reports the top2 distinguishing features (based on FoldChange) for each cluster.
  * This is a P-O-C of what can be done, but avoid over-interpreting results (I purposefully simplified analysis to speed it up).
```{r top DEGs, eval = T, echo = T, fig.width = 11, fig.height=6}
# find markers for every cluster compared to all remaining cells, report only the positive ones

adenocarcino_subset.markers <- FindAllMarkers(adenocarcino_subset,
                                              only.pos = TRUE,
                                              min.pct = 0.5,
                                              max.cells.per.ident = 100,
                                              logfc.threshold = 1)

#give me only the top 2 DEGs for each cluster.
topmarkers = adenocarcino_subset.markers %>% group_by(cluster) %>% slice_max(n = 2, order_by = avg_log2FC) 

#Dot Plot
plot_dot = DotPlot(adenocarcino_subset, features = unique(topmarkers$gene), cols = c("darkblue", "yellow"), dot.scale = 8) + RotatedAxis()
plot_dot

```

# Save
```{r save, timehook = F, message = T, echo = F ,eval = F}
###session
message(paste0('Saving an .RDS object here: ',file.path(params$datapath,params$outputpath,'output')))
dir.create(file.path(params$datapath,params$outputpath,'output'), showWarnings = FALSE)
saveRDS(adenocarcino_subset, file = file.path(params$datapath,params$outputpath,'output/adenocarcino_final.rds'))
```

# session info  
```{r session, timehook = F, message= T}
###session
sessionInfo()
```

