---
title: "scRNA Qc Filters Proof-Of-Concept"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
---

```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/renseb01/Documents/scRNA/scRNA_lung_adeno/data/aggr3018_3019_3027_3028/filtered_feature_bc_matrix')
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE, cache.lazy=FALSE,timehook=TRUE)
```

```{r, requirements,echo=F}
library(dplyr) #data handling
library(Seurat) #scRNA
library(patchwork) #data viz
library(sctransform) #data normalization
library(ggplot2) #data viz
library(DT) #data viz
library(scAnnotatR) #cell type annotation (WARNING: this is currently installed via devtools::install_git('griss/scAnnotatr@develop'))
library(scuttle)
```


# Loading data  
  * `barcodes.tsv.gz`, `features.tsv.gz` and `matrix.mtx.gz` files come directly from  `cell ranger`.
  * Alternatively, we could load each sample separately, but the results are essentially the same (i.e. exact same number of cells, but slightly different count numbers due to the `cellranger aggr` normalization).
```{r initial_load,message = T}
# Load the cellranger dataset
adenocarcino.data <- Read10X(data.dir = ".")

# Initialize the Seurat object with the raw data. 
adenocarcino <- CreateSeuratObject(counts = adenocarcino.data, project = "adenocarcino")

adenocarcino[["percent.mt"]] <- PercentageFeatureSet(adenocarcino, pattern = "^MT-")
```


```{r initial object,message = T,echo = F}
message('Initial seurat object: ')
adenocarcino

```



# Quality control (scater::isOutlier()) 
  * Note that the minimum (lower) filters are a bit silly... 
  * Otherwise, the filters appear good, but IMHO too conservative for `nCount` (UMI) & `nFeature` (genes).
```{r qc,message  = T, fig.width = 11, fig.height=8,echo=F,eval = T}
mt_isOutlier = attr(isOutlier(adenocarcino@meta.data$percent.mt),'thresholds')
nCount_isOutlier = attr(isOutlier(adenocarcino@meta.data$nCount_RNA),'thresholds')
nFeature_isOutlier = attr(isOutlier(adenocarcino@meta.data$nFeature_RNA),'thresholds')

QC_filters = data.frame(rbind(mt_isOutlier,nCount_isOutlier,nFeature_isOutlier))
QC_filters = signif(QC_filters,4)

datatable(QC_filters,caption = htmltools::tags$caption(htmltools::strong("Filters as calculated by isOutlier()"), style="color:darkred"), options = list(dom = 't'))


#filters
# Visualize QC metrics as a violin plot
p1 <- VlnPlot(adenocarcino, features = "nFeature_RNA",  log = T) + geom_hline(yintercept = QC_filters[3,2],col = 'blue',linewidth =2) + theme(legend.position = 'none') 
p2 <- VlnPlot(adenocarcino, features = "nCount_RNA", log = T) + geom_hline(yintercept = QC_filters[2,2],col = 'blue',linewidth =2) + theme(legend.position = 'none')
p3 <- VlnPlot(adenocarcino, features = "percent.mt", log = T) + geom_hline(yintercept = QC_filters[1,2],col = 'blue',linewidth =2) + theme(legend.position = 'none')

wrap_plots(p1, p2, p3, ncol = 3)


adenocarcino_subset <- subset(adenocarcino, subset = 
                                nFeature_RNA < QC_filters[3,2] &
                                nCount_RNA < QC_filters[2,2] &
                                percent.mt < QC_filters[1,2])


message('Filtered seurat object: ')
adenocarcino_subset


```

```{r manual,fig.width = 11,echo =F, fig.height=8,eval = F}
QC_manual = data.frame(rbind(c(250,5000),c(500,25000),c(0,27.65)))
rownames(QC_manual) = c('nFeature (genes)','nCount (UMI)','MT')
colnames(QC_manual) = colnames(QC_filters)
datatable(QC_manual,caption = htmltools::tags$caption(htmltools::strong("Filters as calculated by isOutlier() and then modified manually"), style="color:darkred"), options = list(dom = 't'))

# Visualize QC metrics as a violin plot
p1 <- VlnPlot(adenocarcino, features = "nFeature_RNA",  log = T) + geom_hline(yintercept = QC_manual[1,1],col = 'blue',linewidth =2,linetype='dashed') + geom_hline(yintercept = QC_manual[1,2],col = 'blue',linewidth =2,linetype='dashed')+ theme(legend.position = 'none') 

p2 <- VlnPlot(adenocarcino, features = "nCount_RNA", log = T) + geom_hline(yintercept = QC_manual[2,1],col = 'blue',linewidth =2,linetype='dashed') + geom_hline(yintercept = QC_manual[2,2],col = 'blue',linewidth =2,linetype='dashed')+ theme(legend.position = 'none')

p3 <- VlnPlot(adenocarcino, features = "percent.mt", log = T) + geom_hline(yintercept = QC_manual[3,1],col = 'blue',linewidth =2,linetype=2) + geom_hline(yintercept = QC_manual[3,2],col = 'blue',linewidth =2,linetype='dashed')+ theme(legend.position = 'none')

wrap_plots(p1, p2, p3, ncol = 3)


adenocarcino_subset_manual <- subset(adenocarcino, subset = 
                                nFeature_RNA < QC_manual[1,2] &
                                nCount_RNA < QC_manual[2,2] &
                                percent.mt < QC_manual[3,2] &
                                nFeature_RNA > QC_manual[1,1] &
                                nCount_RNA > QC_manual[2,1] &
                                percent.mt > QC_manual[3,1]
                              )
```


```{r filtered,echo = F, message =T ,eval = F}
message('Filtered seurat object: ')
adenocarcino_subset
```


```{r isoulier per sample,echo = F}
#sample specific IDs are here.
aggr = read.csv('aggregation.csv')
aggr$method = rep(c('all','depleted'),4)
aggr$patient = c(rep(c('P4'),4),rep(c('P6'),4))
aggr$samplename = unlist(sapply( strsplit(aggr$sample_id,split="_"), "[", 1))
aggr$sampletag = 1:8
aggr$type = c('normal','normal','tumor','tumor','normal','normal','tumor','tumor')
aggr$methodXtype = paste0(aggr$method,'_',aggr$type)


#fill the metadata
adenocarcino[["sampletag"]] = unlist(sapply(strsplit(rownames(adenocarcino@meta.data),split="-"), "[", 2))
merged_meta = merge(adenocarcino@meta.data,aggr[,-2],by = "sampletag")
adenocarcino@meta.data[,1:ncol(merged_meta)] = merged_meta
colnames(adenocarcino@meta.data)[1:ncol(merged_meta)] = colnames(merged_meta)


adenocarcino.list <- SplitObject(adenocarcino, split.by = "sample_id")

thresholds = data.frame(sample_id =names(adenocarcino.list),mt_min=0,mt_max=0,nCount_min=0,nCount_max = 0,nFeature_min=0,nFeature_max=0)
for(i in seq_along(adenocarcino.list)){
  mt_isOutlier = attr(isOutlier(adenocarcino.list[[i]]@meta.data$percent.mt),'thresholds')
  nCount_isOutlier = attr(isOutlier(adenocarcino.list[[i]]@meta.data$nCount_RNA),'thresholds')
  nFeature_isOutlier = attr(isOutlier(adenocarcino.list[[i]]@meta.data$nFeature_RNA),'thresholds')
  
  thresholds[i,c(2,3)] = signif(mt_isOutlier,6)
  thresholds[i,c(4,5)] = signif(nCount_isOutlier,6)
  thresholds[i,c(6,7)] = signif(nFeature_isOutlier,6)
  
}

datatable(thresholds)
```

# Filter per sample, then merge object.
```{r filter per sample then merge,echo = F}
adenocarcino.list.subset= adenocarcino.list
for(i in seq_along(adenocarcino.list)){
  
adenocarcino.list.subset[[i]] <- subset(adenocarcino.list[[i]], subset = 
                                nFeature_RNA > thresholds[i,6] &
                                nCount_RNA > thresholds[i,4] &
                                percent.mt > thresholds[i,2] &
                               nFeature_RNA < thresholds[i,7] &
                                nCount_RNA < thresholds[i,5] &
                                percent.mt < thresholds[i,3])


#adenocarcino.list.subset[[i]] <- subset(adenocarcino.list[[i]], subset = 
 #                              nFeature_RNA > 250 &
  #                              nCount_RNA > 500 &
   #                             percent.mt > thresholds[i,2] &
    #                            nFeature_RNA < 5000 &
     #                           nCount_RNA < 25000 &
      #                          percent.mt < thresholds[i,3])

}


#merge datasets back
adenocarcino_m <- merge(adenocarcino.list.subset[[1]],
                       y = adenocarcino.list.subset[-1],
                       project='irsc')

adenocarcino_m
```

```{r beforeafter plots, fig.width = 11, fig.height=8,echo =F}
hist(log(adenocarcino@meta.data$nCount_RNA,10),breaks = seq(2,5.1,by =0.05),col = "darkred",main='nCount_RNA (UMI) filtering',xlab = 'log(Nb of UMIs per cell,10)')
hist(log(adenocarcino_m@meta.data$nCount_RNA,10),breaks = seq(2,5.1,by =0.05),col = "blue",add=T)
legend(x = 3.5,y = 3000,legend = c('before filtering','filtering (thresholds per sample)'),fill = c('darkred','blue')) 

hist(log(adenocarcino@meta.data$nFeature_RNA,10),breaks = seq(1,4,by =0.05),col = "darkred",main='nFeature_RNA (genes) filtering',xlab = 'log(Nb of genes per cell,10)')
hist(log(adenocarcino_m@meta.data$nFeature_RNA,10),breaks = seq(1,4,by =0.05),col = "blue",add=T)
legend(x = 1,y = 4000,legend = c('before filtering','filtering (thresholds per sample)'),fill = c('darkred','blue','yellow'))  

hist(adenocarcino@meta.data$percent.mt,breaks = seq(0,100,by =0.5),col = "darkred",main='MT percent filtering',xlab = 'MT percent')
hist(adenocarcino_m@meta.data$percent.mt,breaks = seq(0,100,by =0.5),col = "blue",add=T)
legend(x = 40,y = 2000,legend = c('before filtering','filtering (thresholds per sample)'),fill = c('darkred','blue')) 
```


```{r sandbox,message  = T,echo = F, eval = F, fig.width = 11, fig.height=8,echo=F}
#filters

hist(adenocarcino@meta.data$percent.mt,breaks = 200,main= 'MT thresholds')
abline(v=mt_qc[1],col= 'red',lwd=4,lty=2)
abline(v=mt_qc[2],col= 'red',lwd=4,lty=2)
abline(v=mt_qc[2],col= 'blue',lwd=4,lty=3)
abline(v=0,col= 'blue',lwd=4,lty=3)
legend(x = max(adenocarcino@meta.data$percent.mt)/2,y = 2000,legend = c('isOutlier thresholds','manual'),fill = c('red','blue'))

hist(adenocarcino@meta.data$nCount_RNA,breaks = 200,main= 'nCount_RNA (UMI) thresholds')
abline(v=ncount_qc[1],col= 'red',lwd=4,lty=2)
abline(v=ncount_qc[2],col= 'red',lwd=4,lty=2)
abline(v=500,col= 'blue',lwd=4,lty=3)
abline(v=ncount_qc[2],col= 'blue',lwd=4,lty=3)
legend(x = max(adenocarcino@meta.data$nCount_RNA)/2,y = 8000,legend = c('isOutlier thresholds','manual'),fill = c('red','blue'))

hist(adenocarcino@meta.data$nFeature,breaks = 200,main= 'nFeature (genes) thresholds')
abline(v=nFeature_qc[1],col= 'red',lwd=4,lty=2)
abline(v=nFeature_qc[2],col= 'red',lwd=4,lty=2)
abline(v=150,col= 'blue',lwd=4,lty=3)
abline(v=nFeature_qc[2],col= 'blue',lwd=4,lty=3)
legend(x = max(adenocarcino@meta.data$nFeature)/2,y = 2000,legend = c('isOutlier thresholds','manual'),fill = c('red','blue'))
```
