---
title: "Single Nuclei vs Single cells (non-depleted) -- Normal tissue, 2 patients"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
params:
  datapath: 'C:/Users/renseb01/Documents/scRNA/scRNA/data'
  outputpath: '../..' 
---

# Setup  
```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = params$datapath)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE, cache.lazy=FALSE)
```

```{r, requirements}
library(Seurat) #scRNA analyses
library(CellChat)
library(patchwork)
library(ComplexHeatmap)
options(stringsAsFactors = FALSE)
```

# Loading data  
```{r loading,message  = T, fig.width = 11, fig.height=8,echo=T}  
seurat.objects =readRDS(file.path(params$datapath,"adeno.24samples_CNV.rds"))
filenames =  list.files(params$datapath,pattern = '-C[ST]T-',full.names = F)

metadata = data.frame()
cells = NULL
for(i in 1:12)
{
  if(i>1) colnames(metadata) = colnames(seurat.objects[[i]]@meta.data) 
  metadata = rbind(metadata,seurat.objects[[i]]@meta.data)
  cells[i] = median(seurat.objects[[i]]@meta.data$nFeature_RNA)
}

metadata$unique = paste(metadata$method,metadata$type,metadata$patient,sep = '___')

nb_cells_per_sample = as.data.frame(table(metadata$unique),stringsAsFactors = F)
```


# get data
```{r cellchat}
combined_permethodN = readRDS(file.path(params$datapath,'combined_permethod_pertypeN24.rds'))
combined_permethodT = readRDS(file.path(params$datapath,'combined_permethod_pertypeT24.rds'))

cellchats = list()
common_cell_types = intersect(unique(combined_permethodN[[1]]@meta.data$predicted.ann_level_3),unique(combined_permethodN[[2]]@meta.data$predicted.ann_level_3))
common_cell_types = common_cell_types[common_cell_types!= 'unknown']


for(i in 1:4){
  if(i < 3) seurat.object = combined_permethodN[[i]]  
  if(i > 2) seurat.object = combined_permethodT[[i-2]]   

data.input = seurat.object@assays$SCT
meta = seurat.object@meta.data # a dataframe with rownames containing cell meta data
cell.use = rownames(meta) # extract the cell names from disease data
#cell.use = rownames(meta)[meta$predicted.ann_level_3 %in% common_cell_types] # extract the cell names from disease data
#cell.use = rownames(meta)[!(meta$predicted.ann_level_3 %in% c("Lymphatic EC differentiating",'unknown','None'))] 


# Prepare input data for CelChat analysis
data.input = data.input[, cell.use]
meta = meta[cell.use, ]

#create cellchat dataset
group.by = 'predicted.ann_level_2'
cellchat <- createCellChat(object = data.input, meta = meta, group.by = group.by)
cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = group.by)

#set-up DB
CellChatDB <- CellChatDB.human #
#dplyr::glimpse(CellChatDB$interaction)

# set the used database in the object
CellChatDB.ecm <- subsetDB(CellChatDB, search = "ECM-Receptor")
CellChatDB.sec <- subsetDB(CellChatDB, search = "Secreted Signaling")
CellChatDB.ccc <- subsetDB(CellChatDB, search = "Cell-Cell Contact")
#CellChatDB.use = CellChatDB
cellchat@DB <- CellChatDB.ccc

#subset
cellchat <- subsetData(cellchat) 

#expresion
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

#compute network
cellchat <- computeCommunProb(cellchat,trim=0.1,type= 'truncatedMean')
cellchat <- filterCommunication(cellchat, min.cells = 10)
cellchat <- aggregateNet(cellchat)


#Compute the communication probability on signaling pathway
cellchat <- computeCommunProbPathway(cellchat)

cellchats[[i]] = cellchat 
print(paste0(i,'---',Sys.time()))
}
# get data
```


# plot data
```{r plots}
#lift ann levels 
group.new = levels(cellchats[[1]]@idents)
cellchats[[2]] = liftCellChat(cellchats[[2]], group.new)
cellchats[[3]] = liftCellChat(cellchats[[3]], group.new)
cellchats[[4]] = liftCellChat(cellchats[[4]], group.new)


mergedcellchat <- mergeCellChat(cellchats, add.names = c('cell_Normal','nucleus_Normal','cell_Tumor','nucleus_Tumor'),cell.prefix = T)
#mergedcellchat <- mergeCellChat(cellchats, add.names = c(1,2,3,4),cell.prefix = T)


#stacked bar shows the strengh of interaction in the four groups, ranked based on prevalence in scRNA-seq normal samples.
stacked2 = rankNet(mergedcellchat, mode = "comparison", stacked = F, do.stat = TRUE,return.data = T,comparison = c(1,2,3,4),color.use = c(RColorBrewer::brewer.pal(n = 6, name = "Paired"))[c(1,2,5,6)])
top10 = stacked2$signaling.contribution %>% group_by(name) %>% summarise(contribution = sum(contribution))
top10 = top10$name[order(top10$contribution,decreasing = T)][1:10]
stacked1 = rankNet(mergedcellchat, mode = "comparison", stacked = T, do.stat = TRUE,return.data = F,signaling=top10, comparison = c(1,2,3,4),color.use = c(RColorBrewer::brewer.pal(n = 6, name = "Paired"))[c(1,2,5,6)]) + theme(legend.position = 'none')
stacked2 = rankNet(mergedcellchat, mode = "comparison", stacked = F, do.stat = TRUE,return.data = F,signaling=top10, comparison = c(1,2,3,4),color.use = c(RColorBrewer::brewer.pal(n = 6, name = "Paired"))[c(1,2,5,6)])

pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat_stacked.pdf'),width = 12,height = 8) 
stacked2+ netAnalysis_contribution(cellchats[[1]], signaling = 'MHC-I') + netAnalysis_contribution(cellchats[[1]], signaling = 'CADM')
dev.off()

#For some genes, like MHC, barplot shows little difference between Normal/tumor, big diff between cell/nuclei in the total strength of interactions. 
#In network again we see little difference between Normal/tumor again big difference between cell-nucleus NORMAL.
CellChatDB.ccc$interaction[CellChatDB.ccc$interaction$pathway_name=='MHC-I',]
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat_MHC.pdf'),width = 12,height = 15) 
par(mfrow = c(2,2), xpd=T)
gene = 'MHC-I'
wmax = getMaxWeight(cellchats, slot.name = c("netP"), attribute = gene)
netVisual_aggregate(cellchats[[1]], signaling = gene, layout = "circle", edge.width.max = 50, signaling.name = paste0(gene,'~cell (Normal)'), edge.weight.max = wmax)
netVisual_aggregate(cellchats[[2]], signaling = gene, layout = "circle", edge.width.max = 50, signaling.name = paste0(gene,'~nucleus (Normal)'), edge.weight.max = wmax)
netVisual_aggregate(cellchats[[3]], signaling = gene, layout = "circle", edge.width.max = 50, signaling.name = paste0(gene,'~cell (Tumor)'), edge.weight.max = wmax)
netVisual_aggregate(cellchats[[4]], signaling = gene, layout = "circle", edge.width.max = 50, signaling.name = paste0(gene,'~nucleus (Tumor)'), edge.weight.max = wmax)
dev.off()

#For other genes like CADM, EPHA and SEMA6, barplot shows big difference in nuclei tumor. When looking at network we see a different network topology in the nuclei tumor, which we would have missed with only cell sequencing. With snRNA-seq it is pretty obvious the normal/tumor difference. The CADM cell-cell interaction pathway is comprised of two self-interacting ligand-receptor genes (CADM1 and CADM3), where CADM1 is involved in the formation and maintenance of epithelial structure and both genes are known as tumor suppressor genes in various cancers, including lung  (https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0082894). . Same story for SEMA6 and EPHA, both known to involved in oncogenesis (ref, ref, ref) 
CellChatDB.ccc$interaction[CellChatDB.ccc$interaction$pathway_name=='CADM',]
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat_CADM.pdf'),width = 12,height = 15) 
par(mfrow = c(2,2), xpd=T)
gene = 'CADM'
wmax = getMaxWeight(cellchats, slot.name = c("netP"), attribute = gene)
netVisual_aggregate(cellchats[[1]], signaling = gene, layout = "circle", edge.width.max = 20, signaling.name = paste0('CADM1~CADM5','~cell (Normal)'), edge.weight.max = wmax)
netVisual_aggregate(cellchats[[2]], signaling = gene, layout = "circle", edge.width.max = 20, signaling.name = paste0('CADM1~CADM5','~nucleus (Normal)'), edge.weight.max = wmax)
netVisual_aggregate(cellchats[[3]], signaling = gene, layout = "circle", edge.width.max = 20, signaling.name = paste0('CADM1~CADM5','~cell (Tumor)'), edge.weight.max = wmax)
netVisual_aggregate(cellchats[[4]], signaling = gene, layout = "circle", edge.width.max = 20, signaling.name = paste0('CADM1~CADM5','~nucleus (Tumor)'), edge.weight.max = wmax)
dev.off()

#
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat_ALL.pdf'),width = 12,height = 15) 
par(mfrow = c(4,2), xpd=T)
netVisual_circle(cellchats[[1]]@net$count, vertex.weight = as.numeric(table(cellchats[[1]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ cell (Normal)")
netVisual_circle(cellchats[[2]]@net$count, vertex.weight = as.numeric(table(cellchats[[2]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ nucleus (Normal)")
netVisual_circle(cellchats[[3]]@net$count, vertex.weight = as.numeric(table(cellchats[[3]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ cell (Tumor)")
netVisual_circle(cellchats[[4]]@net$count, vertex.weight = as.numeric(table(cellchats[[4]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ nucleus (Tumor)")
dev.off()


#comm probabilities
p_com=list()
for(i in 1:4){
  cellchats[[i]] = netAnalysis_computeCentrality(cellchats[[i]])
  p_com[[i]]=netAnalysis_signalingRole_scatter(cellchats[[i]],weight.MinMax= c(0,350))
}

pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat1_com_probs.pdf'),width = 12,height = 6) 
(p_com[[1]]+p_com[[2]])/(p_com[[3]]+p_com[[4]])
dev.off()
#




















































333333

netVisual_aggregate(cellchats[[1]], signaling = 'APP', layout = "circle", edge.width.max = 3, signaling.name = 'APP~cell (Normal)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'APP'))
netVisual_aggregate(cellchats[[2]], signaling = 'APP', layout = "circle", edge.width.max = 3, signaling.name = 'APP~nucleus (Normal)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'APP'))
netVisual_aggregate(cellchats[[3]], signaling = 'APP', layout = "circle", edge.width.max = 3, signaling.name = 'APP~cell (Tumor)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'APP'))
netVisual_aggregate(cellchats[[4]], signaling = 'APP', layout = "circle", edge.width.max = 3, signaling.name = 'APP~nucleus (Tumor)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'APP'))

netVisual_aggregate(cellchats[[1]], signaling = 'PTPRM', layout = "circle", edge.width.max = 3, signaling.name = 'APP~cell (Normal)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'PTPRM'))
netVisual_aggregate(cellchats[[2]], signaling = 'PTPRM', layout = "circle", edge.width.max = 3, signaling.name = 'APP~nucleus (Normal)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'PTPRM'))
netVisual_aggregate(cellchats[[3]], signaling = 'PTPRM', layout = "circle", edge.width.max = 3, signaling.name = 'APP~cell (Tumor)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'PTPRM'))
netVisual_aggregate(cellchats[[4]], signaling = 'PTPRM', layout = "circle", edge.width.max = 3, signaling.name = 'APP~nucleus (umor)', edge.weight.max = getMaxWeight(cellchats, slot.name = c("netP"), attribute = 'PTPRM'))
#par(mfrow = c(1,2), xpd=TRUE)
#draw(p7 + p8, ht_gap = unit(0.5, "cm"))




#mergedcellchat_lev2 <- mergeCellChat(cellchats[c(3,4)], add.names =c('cell_NORMAL','cell_TUMOR'),cell.prefix = T)




#nb of interactions
p4=compareInteractions(mergedcellchat_normal_CellNucleus, show.legend = F,group = c(1,2))
p5=compareInteractions(mergedcellchat_lev2, show.legend = F,measure = 'weight',group = c(1,2))


#compare 
p6_temp <- rankNet(mergedcellchat, mode = "comparison", stacked = T, do.stat = TRUE,return.data = T,comparison = c(1,2,3,4))


#new order
data = p6_temp$signaling.contribution
p.adj = p.adjust(data$pvalues)
data =data[p.adj<0.05,]
pathway.signif = data[rownames(data)==data$name,]
pathway.signif = rownames(pathway.signif)[order(pathway.signif$contribution,decreasing=F)]
data$name =factor(data$name, levels = pathway.signif)
data$group = factor(data$group,levels = c('cell_NORMAL','cell_TUMOR'))


#
p6 = ggplot(data, aes(fill=group, y=name, x=contribution.scaled)) + 
    geom_bar(position="fill", stat="identity")+ xlab('Relative contribution\nto signal')

#heat
cellchats[[3]] = netAnalysis_computeCentrality(cellchats[[3]])
cellchats[[4]] = netAnalysis_computeCentrality(cellchats[[4]])
pathway.union <- union(cellchats[[3]]@netP$pathways,cellchats[[4]]@netP$pathways)
#pathway.signif_reverse = pathway.signif[order(pathway.signif,decreasing=T)]
p7 = netAnalysis_signalingRole_heatmap(cellchats[[3]], pattern = "all", signaling = rev(levels(data$name)), title = 'normal', width = 5, height = 15, color.heatmap = "OrRd")
p8 = netAnalysis_signalingRole_heatmap(cellchats[[4]], pattern = "all", signaling = rev(levels(data$name)), title = 'tumor', width = 5, height = 15, color.heatmap = "OrRd")



pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat1.pdf'),width = 12,height = 6) 
par(mfrow = c(1,3), xpd=TRUE)
#networks
p1= netVisual_circle(cellchats[[3]]@net$count, vertex.weight = as.numeric(table(cellchats[[3]]@idents)), weight.scale = T, label.edge= F, title.name = "Normal cell")
p2= netVisual_circle(cellchats[[4]]@net$count, vertex.weight = as.numeric(table(cellchats[[4]]@idents)), weight.scale = T, label.edge= F, title.name = "Tumor cell")
p3= netVisual_diffInteraction(mergedcellchat_lev2, weight.scale = T,comparison=c('cell_NORMAL','cell_TUMOR'),title= 'Normal minus Tumor Interactions')
dev.off()

#
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat2.pdf'),width = 8,height = 6) 
p4+p5+p6
dev.off()

#
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat3.pdf'),width = 12,height = 15) 
#par(mfrow = c(1,2), xpd=TRUE)
draw(p7 + p8, ht_gap = unit(0.5, "cm"))
dev.off()



































###1
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat1.pdf'),width = 12,height = 6) 
par(mfrow = c(1,3), xpd=TRUE)
#cellN
netVisual_circle(cellchats[[1]]@net$count, vertex.weight = as.numeric(table(cellchats[[1]]@idents)), weight.scale = T, label.edge= F, title.name = "Normal cell")
#cellT
netVisual_circle(cellchats[[2]]@net$count, vertex.weight = as.numeric(table(cellchats[[2]]@idents)), weight.scale = T, label.edge= F, title.name = "Tumor cell")
#N-T differences
netVisual_diffInteraction(mergedcellchat, weight.scale = T,comparison=c('cell_NORMAL','cell_TUMOR'),title= 'Normal minus Tumor Interactions')
#diff number
dev.off()

###2
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat2.pdf'),width = 12,height = 6) 
plot_comp = compareInteractions(mergedcellchat, show.legend = F,group = c(1,2))
#diff weight
plot_comp_w = compareInteractions(mergedcellchat, show.legend = F,measure = 'weight',group = c(1,2))
#inter
plot_rank = rankNet(mergedcellchat, mode = "comparison", stacked = T, do.stat = T,comparison = c(1,2))
plot_rank_W = rankNet(mergedcellchat, mode = "comparison", stacked = F, do.stat = T,comparison = c(1,2))
plot_comp | plot_comp_w | plot_rank | plot_rank_W
dev.off()

###3 
pdf(file.path(params$datapath,'../results/Figures_July2023/cellchat3.pdf'),width = 12,height = 6) 
par(mfrow = c(1,2), xpd=TRUE)
netVisual_aggregate(cellchats[[1]], signaling = 'MHC-II', layout = "circle",signaling.name= 'MHC-II ~ cell')
netVisual_aggregate(cellchats[[3]], signaling = 'MHC-II', layout = "circle",signaling.name= 'MHC-II ~ tumor')
dev.off()






# show all the significant interactions (L-R pairs) associated with certain signaling pathways
# show all the significant signaling pathways from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_chord_gene(cellchats[[1]], sources.use = c(1,2), targets.use = c(3), slot.name = "netP", legend.pos.x = 10)









###plot all the pathwayws HERE yeah!!!!
x = netVisual_aggregate(cellchats[[1]], signaling = 'NOTCH', layout = "circle")
x = netVisual_aggregate(cellchats[[2]], signaling = 'NOTCH', layout = "circle")



p1 = compareInteractions(mergedcellchat, show.legend = F, group = c(1,2))
p2 = compareInteractions(mergedcellchat, show.legend = F, group = c(1,2),measure = 'weight')

par(mfrow = c(1,2), xpd=TRUE)
p3=netVisual_diffInteraction(mergedcellchat, weight.scale = T)
p4=netVisual_diffInteraction(mergedcellchat, weight.scale = T, measure = "weight")


p5 <- netVisual_heatmap(mergedcellchat)
#> Do heatmap based on a merged object
p6 <- netVisual_heatmap(mergedcellchat, measure = "weight")
#> Do heatmap based on a merged object
p5 + p6


#compare datasets
gg1 <- rankNet(mergedcellchat, mode = "comparison", stacked = T, do.stat = TRUE)
gg2 <- rankNet(mergedcellchat, mode = "comparison", stacked = F, do.stat = TRUE)
gg1 + gg2

```
# real data  
```{r cellchat}
# Here we load a scRNA-seq data matrix and its associated cell meta data
cellchats = list()

types = c('Epithelial','Immune')
tissue = c('normal','tumor')
for(i in c(1,2)){ #normal, tumor
  for(j in seq_along(types)){
    
#data.input = GetAssayData(cellchat_dataset[[i]]) # normalized data matrix
if(i == 1) seurat.object = combined_permethodN[[1]]    
if(i == 2) seurat.object = combined_permethodT[[1]]     

#seurat.object = seurat.objects[[10]]

data.input = seurat.object@assays$SCT
meta = seurat.object@meta.data # a dataframe with rownames containing cell mata data
cell.use = rownames(meta) # extract the cell names from disease data
#cell.use = rownames(meta)[meta$predicted.ann_level_1 == types[j]] # extract the cell names from disease data


# Prepare input data for CelChat analysis
data.input = data.input[, cell.use]
meta = meta[cell.use, ]

meta$seurat_clusters2 = paste('cluster',meta$seurat_clusters)

#meta$seurat_clusters2[meta$seurat_clusters=='7'] = 'c7_tumor_epi'

#create cellchat dataset
group.by = 'predicted.ann_level_2'
cellchat <- createCellChat(object = data.input, meta = meta, group.by = group.by)
cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = group.by)

#set-up DB
CellChatDB <- CellChatDB.human #
#dplyr::glimpse(CellChatDB$interaction)

# set the used database in the object
#CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling")
CellChatDB.use <- subsetDB(CellChatDB, search = "Cell-Cell Contact")
#CellChatDB.use = CellChatDB
cellchat@DB <- CellChatDB.use

#subset
cellchat <- subsetData(cellchat) 
#future::plan("multiprocess", workers = 4) # do parallel

#expresion
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

#compute network
cellchat <- computeCommunProb(cellchat,trim=0.1,type= 'truncatedMean')
cellchat <- filterCommunication(cellchat, min.cells = 10)
cellchat <- aggregateNet(cellchat)

cellchats[[(i*2)-2 + j]] = cellchat

print(paste0('Done sample ',tissue[i], ' for ',types[j],' cells, Time is:',Sys.time()))
}
}
```


par(mfrow = c(1,2)) 
a = netVisual_circle(cellchat_nuc@net$count, vertex.weight = as.numeric(table(cellchat_nuc@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ Normal epithelial")
b = netVisual_circle(cellchat_cell@net$count, vertex.weight = as.numeric(table(cellchat_cell@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ Normal epithelial")


```{r plots}
#plot results
par(mfrow = c(2,2)) 
a=netVisual_circle(cellchats[[1]]@net$count, vertex.weight = as.numeric(table(cellchats[[1]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ Normal epithelial")
b= netVisual_circle(cellchats[[3]]@net$count, vertex.weight = as.numeric(table(cellchats[[3]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ Tumor epithelial")
c=netVisual_circle(cellchats[[2]]@net$count, vertex.weight = as.numeric(table(cellchats[[2]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ Normal Immune")
d=netVisual_circle(cellchats[[4]]@net$count, vertex.weight = as.numeric(table(cellchats[[4]]@idents)), weight.scale = T, label.edge= F, title.name = "Number of interactions ~ Normal Immune")

#groupSize <- as.numeric(table(cellchat@idents))
pdf(file.path(params$datapath,'../results/FigX_ligand_receptor.pdf'),width = 12,height = 7) 

#a;b;c;d
dev.off()
#netVisual_aggregate(cellchat, signaling = "CXCL", layout = "chord")

```










```{r plots}
for(i in 1:12){
  data = GetAssayData(seurat.objects[[i]],slot = 'data')
  meta = seurat.objects[[i]]@meta.data
  rm = rowMeans(data)
  print(head(data[order(rm,decreasing=T),1:5],5))
}



seurat.object = combined_permethodN[[1]] 
data = seurat.object@assays$RNA
#data = GetAssayData(seurat.object,slot = 'data')
meta = seurat.object@meta.data
cd274  = data[rownames(data)== 'CD274',]
cd274_expressed = meta$seurat_clusters[as.numeric(cd274)>0]
table(cd274_expressed)/table(meta$seurat_clusters)




malat1  = data[rownames(data)== 'MALAT1',]
meta$malat1 = malat1[1,]

combined_permethodN[[1]]@meta.data$malat1 = malat1[1,]
data_summarised = meta %>% 
      group_by(seurat_clusters) %>%
      summarise(mea = mean(malat1),med = median(malat1),.groups='keep')

print(data_summarised,n=22)

```






