---
title: "Single Cells - 20 patients"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
params:
  datapath: 'C:/Users/renseb01/Documents/scRNA/scRNA/data'
  outputpath: '../..' 
---

# Setup  
* This is a scRNA project 
* Note that you can create this document by running:  
  * `Rscript -e "rmarkdown::render('scRNA_sannotation_all.RMd',params = list())"` and `knit`-it directly in `RStudio`.
  
```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = params$datapath)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE, cache.lazy=FALSE)
```

```{r, requirements}
library(Seurat) #scRNA analyses
library(sctransform) #data normalization
library(scater) #outlier values predictions
library(DoubletFinder) #find doublets? 
library(dplyr) #data handling
library(patchwork) #data viz
library(ggplot2) #data viz
library(DT) #data viz
library(plotly)
source('../R/doubletfinder.R')
source('../R/azimuth_analysis.R')
```

# Loading data  
  * `barcodes.tsv.gz`, `features.tsv.gz` and `matrix.mtx.gz` files come directly from  `cell ranger`.


```{r filenames,message  = T, fig.width = 11, fig.height=8,echo=T}  
filenames =  list.files('.',pattern = '-C[ST]T-',full.names = F)
filepaths = file.path(getwd(),filenames,'filtered_feature_bc_matrix')

filenames
```


```{r qc,message  = T, fig.width = 11, fig.height=8,echo=F}  
#create seurat objects
seurat.objects = list()
l = length(filenames)

#create filters objects
QC_filters=data.frame(nCount_min=rep(0,l),nCount_max=rep(0,l),
                    nFeature_RNA_min=rep(0,l),nFeature_RNA_max=rep(0,l),
                    percent.mt_min=rep(0,l),percent.mt_max=rep(0,l))


for(i in 1:length(filenames)) {
#for(i in c(1:4,17:20)) {
  # Read 10x data
  cts <- Read10X(data.dir = filepaths[i])
  seurat.objects[[i]] = CreateSeuratObject(counts = cts)
  print(seurat.objects[[i]])
  seurat.objects[[i]][["percent.mt"]] <- PercentageFeatureSet(seurat.objects[[i]], pattern = "^MT-")

  # Filters
  QC_filters[i,1:2] = attr(isOutlier(seurat.objects[[i]]@meta.data$nCount),'thresholds')
  QC_filters[i,3:4] = attr(isOutlier(seurat.objects[[i]]@meta.data$nFeature_RNA),'thresholds')
  QC_filters[i,5:6] = attr(isOutlier(seurat.objects[[i]]@meta.data$percent.mt),'thresholds')
  
  # Subset
  seurat.objects[[i]] <- subset(seurat.objects[[i]], subset = nCount_RNA <  QC_filters[i,2])
  seurat.objects[[i]] <- subset(seurat.objects[[i]], subset = nFeature_RNA <  QC_filters[i,4])
  seurat.objects[[i]] <- subset(seurat.objects[[i]], subset = percent.mt <  ifelse(QC_filters[i,6]>25,25,QC_filters[i,6]))

  # Generate metadata
  
  # Generate metadata
  temp = strsplit(filenames[i],split = '_')[[1]]
  if(temp[2]=='all')temp[2]='Cell'
  if(temp[2]=='depletion')temp[2]='immune depleted cells'
  if(temp[2]=='SN')temp[2]='Nucleus'
  meta = c(filenames[i],temp[2],ifelse(regexpr('CST',temp[1])>0,'Normal','Tumor'),strsplit(temp[1],fixed = T,'-')[[1]][2])
  seurat.objects[[i]]@meta.data$samplename = temp[1]
  seurat.objects[[i]]@meta.data$method = temp[2]
  seurat.objects[[i]]@meta.data$type = ifelse(regexpr('CST',temp[1])>0,'Normal','Tumor')
  seurat.objects[[i]]@meta.data$patient = strsplit(temp[1],'-')[[1]][1]
  
  # Normalisation
  seurat.objects[[i]] <- SCTransform(seurat.objects[[i]], method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = F)
  
  print(paste0(i,' --- Time: ',Sys.time(),' -- seurat object ready (subsetting, normalisation)'))
}
names(seurat.objects) =  filenames
```


# Doublet Finder
  * find and remove doublet (assuming a 5% doublet rate).
```{r doublet finder, echo=T, message=FALSE, warning=FALSE,include = FALSE, timehook=T}
#precomputed pk's for doubletfinder
#pK_empirical_in = c(0.05,0.29,0.09,0.01, 0.030,0.010,0.020,0.300,0.300,0.001,0.210,0.060,0.010,0.090,0.240,0.040,0.2,0.19,0.24,0.05)
pK_empirical_in = c(0.05,0.29,0.04,0.09,0.01,0.16,0.030,0.010,0.020,0.300,0.300,0.001,0.210,0.060,0.010,0.090,0.240,0.040,0.2,0.19,0.03,0.24,0.05,0.02)
pK_empirical_out =rep(0,length(filenames))
for(i in 1:length(filenames)) {
  doubletfinder.out = doubletfinder_wrapper(seurat.objects[[i]], pK_empirical = pK_empirical_in[i])
  seurat.objects[[i]] = doubletfinder.out[[1]]
  pK_empirical_out[i] =  doubletfinder.out[[2]]
}
```

```{r results, message=T}
summary_table = data.frame(sample = filenames,
                           type = '' ,
                           method = '',
                           patient = '',
                           nb_cells = '',
                           nb_genes = '')

for(i in seq_along(filenames)){
  summary_table$type[i] = unique(seurat.objects[[i]]$type)
  summary_table$method[i] = unique(seurat.objects[[i]]$method)
  summary_table$patient[i] =  unique(seurat.objects[[i]]$patient)
  summary_table$nb_cells[i] =  ncol(seurat.objects[[i]])
   summary_table$nb_genes[i] =  nrow(seurat.objects[[i]])
}

summary_table = summary_table[order(summary_table$type,summary_table$method),]

write.table(summary_table,file.path(params$datapath,"../results/summary_table.csv"),row.names = F, quote = F,sep = '\t')

```



# Filters defined with isOutlier()
```{r DTfilter,echo =F, message = T} 
rownames(QC_filters) = filenames
QC_filters = signif(QC_filters,4)
datatable(QC_filters,caption = htmltools::tags$caption(htmltools::strong("Filters as calculated by isOutlier()"), style="color:darkred"))
```


# SAVE DATA  
```{r SAVE DATA,message  = T, fig.width = 11, fig.height=8,echo=T}  

saveRDS(seurat.objects, file = file.path(params$datapath,"adeno.24samples_27_07_23.rds"))
```

# DONE DATA  
```{r done,message  = T, fig.width = 11, fig.height=8,echo=T}  
print('toto')

```


