---
title: "Single Cells - 2 Mus musculus samples"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
params:
  datapath: 'C:/Users/renseb01/Documents/joseph_vincent/data'
---

# Setup  
* This is a scRNA project 
```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = params$datapath)
knitr::opts_chunk$set(echo=TRUE, message=TRUE, warning=TRUE, cache=FALSE, cache.lazy=FALSE)
```

# Requirements
```{r, requirements}
library(Seurat) #scRNA analyses
library(SeuratDisk) #scRNA analyses
library(sctransform) #data normalization
library(scater) #outlier values predictions
library(DoubletFinder) #find doublets? 
library(dplyr) #data handling
library(patchwork) #data viz
library(ggplot2) #data viz
library(DT) #data viz
library(plotly)
source('../../scRNA/scRNA/R/doubletfinder.R')
```

# Loading data  
  * `barcodes.tsv.gz`, `features.tsv.gz` and `matrix.mtx.gz` files come directly from  `cell ranger`.
```{r filenames,message  = T, fig.width = 11, fig.height=8,echo=T}  
filenames =  list.files('.',pattern = 'mm10',full.names = F)
filepaths = file.path(getwd(),filenames,'filtered_feature_bc_matrix')
filenames
```

# Load cell rangers datasets and do the initial filtering
```{r qc,message  = T, fig.width = 11, fig.height=8,echo=F}  
#create seurat objects
seurat.objects = list()
l = length(filenames)

#create filters matrix
QC_filters=data.frame(nCount_min=rep(0,l),nCount_max=rep(0,l),
                    nFeature_RNA_min=rep(0,l),nFeature_RNA_max=rep(0,l),
                    percent.mt_min=rep(0,l),percent.mt_max=rep(0,l))


for(i in 1:length(filenames)) {
  # Read 10x data
  cts <- Read10X(data.dir = filepaths[i])
  seurat.objects[[i]] = CreateSeuratObject(counts = cts)
  
  print(seurat.objects[[i]])
  seurat.objects[[i]][["percent.mt"]] <- PercentageFeatureSet(seurat.objects[[i]], pattern = "^mt-")

  # Filters
  QC_filters[i,1:2] = attr(isOutlier(seurat.objects[[i]]@meta.data$nCount),'thresholds')
  QC_filters[i,3:4] = attr(isOutlier(seurat.objects[[i]]@meta.data$nFeature_RNA),'thresholds')
  QC_filters[i,5:6] = attr(isOutlier(seurat.objects[[i]]@meta.data$percent.mt),'thresholds')
  
  # Generate metadata
  seurat.objects[[i]]@meta.data$samplename = strsplit(filenames[i],split = '_')[[1]][1]
  seurat.objects[[i]]@meta.data$filename = filenames[i]

  # Normalisation using SCT
   seurat.objects[[i]] <- SCTransform(seurat.objects[[i]], method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = F)
  
  print(paste0(i,' --- Time: ',Sys.time(),' -- seurat object ready (subsetting, normalisation)'))
}
names(seurat.objects) =  filenames
```

# Merge raw matrices (for violon plots)
```{r mergeraw}
seurat.objects[[3]] =  merge(seurat.objects[[1]],seurat.objects[[2]])
viol_plots = VlnPlot(seurat.objects[[3]], features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3,group.by='samplename')

#Then subset based on QC filters defined above
for(i in 1:2){
  # Subset
  seurat.objects[[i]] <- subset(seurat.objects[[i]], subset = nCount_RNA <  QC_filters[i,2])
  seurat.objects[[i]] <- subset(seurat.objects[[i]], subset = nFeature_RNA <  QC_filters[i,4])
  seurat.objects[[i]] <- subset(seurat.objects[[i]], subset = percent.mt <  QC_filters[i,6])
}

#show filters defined by isOutlier()
rownames(QC_filters) = filenames
QC_filters = signif(QC_filters,4)
datatable(QC_filters,caption = htmltools::tags$caption(htmltools::strong("Filters as calculated by isOutlier()"), style="color:darkred"))
```

# Doublet Finder
  * find and remove doublet (assuming a 5% doublet rate).
  * THis relies on `doubletfinder_wrapper()` which I wrote (in doubletfinder.R). It is a simple wrapper for several functions in `DoubletFinder`. 
```{r doublet finder, echo=T, message=FALSE, warning=FALSE,include = FALSE, timehook=T}
#precomputed pk's for doubletfinder
pK_empirical_in = c(0.17,0.1)
pK_empirical_out =rep(0,length(filenames))
for(i in 1:2) {
  doubletfinder.out = doubletfinder_wrapper(seurat.objects[[i]], pK_empirical = pK_empirical_in[i])
 #doubletfinder.out = doubletfinder_wrapper(seurat.objects[[i]])
  seurat.objects[[i]] = doubletfinder.out[[1]]
  pK_empirical_out[i] =  doubletfinder.out[[2]]
  colnames(seurat.objects[[i]]@meta.data)[regexpr('pANN',colnames(seurat.objects[[i]]@meta.data))>0] = 'pANN_doublet_finder'
}
```

# Save filtered data as .h5Seurat objects
```{r save, echo = T}

#create an temporary object of class assay instead of assay5 for Azimuth (Seurat v4) annotation
for(i in 1:3) {
  temp_v3 <- CreateAssayObject(counts = seurat.objects[[i]][['RNA']]$counts)
  seurat.objects[[i]][["RNA"]] <- temp_v3
}


saveRDS(seurat.objects[[1]],file.path(params$datapath,'CIH_mm10_vM23/CIH_mm10_vM23.rds'))
saveRDS(seurat.objects[[2]],file.path(params$datapath,'NX_mm10_vM23/NX_mm10_vM23.rds'))
saveRDS(seurat.objects, file = file.path(params$datapath,"mm_2samples.rds"))
```


# Annotation 
  * I'm not actually doing it here, because I am running it directly [here](https://app.lungmap.net/app/azimuth-mouse-lung-cellref-seed). I use either `CIH_mm10_vM23.h5Seurat` or `NX_mm10_vM23.h5Seurat` and generate the annotations from there. Once this is done, I save the `azimuth_results.Rds` objects, and load it back below.  

```{r get annotations here}
annotations_CIH = readRDS(file.path(params$datapath,'CIH/azimuth_results.Rds'))#these are annotations downloaded directly from the webserver.
annotations_NX = readRDS(file.path(params$datapath,'NX/azimuth_results.Rds')) #these are annotations downloaded directly from the webserver.

#add metadata
seurat.objects[[1]] <- AddMetaData(object = seurat.objects[[1]],metadata = annotations_CIH$pred.df)
seurat.objects[[2]] <- AddMetaData(object = seurat.objects[[2]],metadata = annotations_NX$pred.df)

#add reference UMAPs
seurat.objects[[1]]@reductions[['proj.umap']] <- CreateDimReducObject(embeddings = annotations_CIH$umap@cell.embeddings, key = 'proj.umap', assay = 'RNA')
seurat.objects[[2]]@reductions[['proj.umap']] <- CreateDimReducObject(embeddings = annotations_NX$umap@cell.embeddings, key = 'proj.umap', assay = 'RNA')

#merge object (remember seurat.objects[[3]] is already the raw data)
seurat.objects[[4]] =  merge(seurat.objects[[1]],seurat.objects[[2]])

#annotations
write.csv(seurat.objects[[4]]@meta.data,file.path(params$datapath,'metadata_mus_musculus_annotations.csv'))

#Expression data (normalised)
data = GetAssayData(seurat.objects[[4]])
data = round(data,6)
write.csv(data,file.path(params$datapath,'assaydata_vst.csv'))

saveRDS(seurat.objects, file = file.path(params$datapath,"mm_2samples.rds"))
write.csv(seurat.objects[[4]]@meta.data,file.path(params$datapath,'metadata_mm.csv'))
```

```{r plots, eval = T}
#dimplots
dimplot_lev1 = DimPlot(seurat.objects[[4]],reduction = 'proj.umap',group.by = 'predicted.lineage_level1',label = T, repel = T, order = T,pt.size = 0.5) + theme(axis.title = element_text(size = 8),legend.position = "none") + ggtitle('Annotation level 1 ~ lung tissue\nMus musculus (scRNA)')

dimplot_lev3 = DimPlot(seurat.objects[[4]],reduction = 'proj.umap',group.by = 'predicted.celltype_level3',label = T, repel = T, order = T,pt.size = 0.5) + theme(axis.title = element_text(size = 8),legend.position = "none") + ggtitle('Annotation level 3 ~ lung tissue\nMus musculus (scRNA)')

dimplot_sam = DimPlot(seurat.objects[[4]],reduction = 'proj.umap',group.by = 'samplename',label = F, repel = T, order = T,pt.size = 0.5) + theme(axis.title = element_text(size = 8),legend.position = "right") + ggtitle('samplename ~ lung tissue\nMus musculus (scRNA)')


#Below are human genes. We can only plot those, if we use human data, so we won't for now...
#hormonal_receptors = c('AR','SLC39A9', 'CACNA1C', 'ESR1', 'ESR2', 'GPER1', 'PGR', 'PGRMC1', 'PGRMC2', 'PAQR5', 'PAQR6', 'PAQR7', 'PAQR8', 'HIF1A', 'EPAS1', 'NFE2L2','PAQR9')
#'PAQR9'
#featplot = FeaturePlot(seurat.objects[[3]],features = hormonal_receptors,order = T)

#pdf(file.path('joseph_vincent/results/featplot_2samples.pdf'),width = 12,height = 12) 
#featplot 
#dev.off()

pdf(file.path(params$datapath,'../results/UMAP_2samples.pdf'),width = 12,height = 12) 
(dimplot_lev1 | dimplot_lev3) / (dimplot_sam  | viol_plots)
dev.off()
  
```



# session info  
```{r session, message= T}
###session
sessionInfo()
```

