---
title: "Single Nuclei vs Single cells (non-depleted) -- Normal tissue, 2 patients"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
params:
  datapath: 'C:/Users/renseb01/Documents/scRNA/scRNA/data'
  outputpath: '../..' 
---

# Setup  
```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = params$datapath)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE, cache.lazy=FALSE)
```

```{r, requirements}
library(Seurat) #scRNA analyses
library(sctransform) #data normalization
library(scater) #outlier values predictions
library(scAnnotatR) #cell type annotation (devtools::install_github('grisslab/scAnnotatR'))
library(DoubletFinder) #find doublets? 
library(RColorBrewer)
library(vegan) #diversity indices

library(dplyr) #data handling
library(patchwork) #data viz
library(ggplot2) #data viz
library(DT) #data viz
library(plotly)
source('../R/doubletfinder.R')
source('../R/azimuth_analysis.R')
```

# Loading data  
```{r loading,message  = T, fig.width = 11, fig.height=8,echo=T}  
seurat.objects =readRDS(file.path(params$datapath,"adeno.12samples.rds"))
filenames =  list.files(params$datapath,pattern = '-C[ST]T-',full.names = F)

metadata = data.frame()
cells = NULL
for(i in 1:12)
{
  if(i>1) colnames(metadata) = colnames(seurat.objects[[i]]@meta.data) 
  metadata = rbind(metadata,seurat.objects[[i]]@meta.data)
  cells[i] = median(seurat.objects[[i]]@meta.data$nFeature_RNA)
}

metadata$unique = paste(metadata$method,metadata$type,metadata$patient,sep = '___')

nb_cells_per_sample = as.data.frame(table(metadata$unique),stringsAsFactors = F)
```

# Integration (12 samples)
  * Here, we follow the [most up-to-date approach of Seurat](https://satijalab.org/seurat/articles/integration_introduction.html#performing-integration-on-datasets-normalized-with-sctransform-1) to integrate different single-cells samples together.
  * This is a computationally expensive step that scales greater than linearly with dataset size, so proceed with caution.
```{r integration, echo = T, message = F}
if(file.exists(file.path(params$datapath,'adeno.combined.sct.rds')) == F){
  features <- SelectIntegrationFeatures(object.list = seurat.objects, nfeatures = 2000)
  seurat.objects <- PrepSCTIntegration(object.list = seurat.objects, anchor.features = features)
  adeno.anchors <- FindIntegrationAnchors(object.list = seurat.objects, normalization.method = "SCT",anchor.features = features)
  adeno.combined.sct <- IntegrateData(anchorset = adeno.anchors, normalization.method = "SCT")
  npc = 30
  message(paste0('Keeping the first ',npc, ' Principal Components for UMAP'))
  adeno.combined.sct <- RunPCA(adeno.combined.sct, verbose = FALSE)
  adeno.combined.sct <- RunUMAP(adeno.combined.sct, reduction = "pca", dims = 1:npc, verbose = FALSE)
  saveRDS(adeno.combined.sct, file = file.path(params$datapath,"adeno.combined.sct.rds"))
} else adeno.combined.sct = readRDS(file.path(params$datapath,'adeno.combined.sct.rds'))
```

# General Plots
```{r UMAPplots}
p1 <- DimPlot(adeno.combined.sct, reduction = "umap", group.by = "patient")+ theme(text = element_text(size = 30))  
p1$labels$colour = 'Patients'
p2 <- DimPlot(adeno.combined.sct, reduction = "umap", group.by = "type", cols = c('yellow','darkblue'))+ theme(text = element_text(size = 30))  
p2$labels$colour = 'Type'
p3 <- DimPlot(adeno.combined.sct, reduction = "umap", group.by = "method")+ theme(text = element_text(size = 30))  
p3$labels$colour = 'Method'
p4 <- DimPlot(adeno.combined.sct, reduction = "umap", group.by = "predicted.ann_level_1", cols = 'Set1') + ggtitle("Cell Type Annotation")+ theme(text = element_text(size = 30))  
p4$labels$colour = 'Annotations'

png(file.path(params$datapath,'../results/Figure4_UMAP_general.png'),width = 2600,height = 1400)
p1 + p2 + p3 + p4 + plot_annotation(tag_levels = 'A')
dev.off()
```


# 'SingleCells','SingleNuclei','Depleted' specific plots
```{r UMAPplots}
if(file.exists(file.path(params$datapath,'combined_permethod.rds')) == F){
npc=30
combined_permethod = list()
for(i in 1:3){
  if((i == 1)&(t==1)) samples = c(1:length(filenames))[regexpr('CST-...._all_',filenames)>0]
  if((i == 2)&(t==1)) samples = c(1:length(filenames))[regexpr('CST-...._SN_',filenames)>0]
  if((i == 3)&(t==1)) samples = c(1:length(filenames))[regexpr('CST-...._depletion_',filenames)>0]
  
  features <- SelectIntegrationFeatures(object.list = seurat.objects[samples], nfeatures = 2000)
  seurat.objects.method <- PrepSCTIntegration(object.list = seurat.objects[samples], anchor.features = features)
  adeno.anchors <- FindIntegrationAnchors(object.list = seurat.objects.method, normalization.method = "SCT", anchor.features = features)
  combined_permethod[[i]] <- IntegrateData(anchorset = adeno.anchors, normalization.method = "SCT")
  combined_permethod[[i]] <- RunPCA(combined_permethod[[i]], verbose = FALSE)
  combined_permethod[[i]] <- RunUMAP(combined_permethod[[i]], reduction = "pca", dims = 1:npc, verbose = FALSE)
  combined_permethod[[i]]@meta.data$level1_type = paste(substring(combined_permethod[[i]]@meta.data$predicted.ann_level_1,1,4),combined_permethod[[i]]@meta.data$type,sep = '_')
  
  combined_permethod[[i]] <- FindNeighbors(combined_permethod[[i]], reduction = "pca", dims = 1:30)
  combined_permethod[[i]] <- FindClusters(combined_permethod[[i]], resolution = 0.5)
}
}
saveRDS(combined_permethod,file.path(params$datapath,"combined_permethod_pertype.rds"))

```

# plots
```{r plots}
###plots
for(i in 1:3){
  methods = c('SingleCells','SingleNuclei','Depleted')
  figure = c(5,7,6)
  pB=DimPlot(combined_permethod[[i]], reduction = "umap", group.by = "level1_type",cols = brewer.pal(n = 10, name = "Paired")[c(5,6,1,2,3,4,9,10,7,8)],shape.by = 'type',pt.size=0.4) + ggtitle('Cells Annotations') + guides(shape='none') + theme(legend.text=element_text(size=8))
    
  pB$labels$colour = 'Annotations'  

  pA = DimPlot(combined_permethod[[i]], reduction = "umap",label = T,group.by =  'integrated_snn_res.0.5') + ggtitle('Clustering') + theme(legend.position = "none")

  ###data creation
  data = as.data.frame(table(combined_permethod[[i]]@meta.data$type,combined_permethod[[i]]@meta.data$predicted.ann_level_1,combined_permethod[[i]]@meta.data$integrated_snn_res.0.5))
  colnames(data)=c('type','Annotation','Clusters','Number of Cells')
  
  data$`Proportions of Cells`= 0
  
  data$`Proportions of Cells`[data$type=='Normal'] = data$`Number of Cells`[data$type=='Normal']/sum(data$`Number of Cells`[data$type=='Normal'])
  
  data$`Proportions of Cells`[data$type=='Tumor'] = data$`Number of Cells`[data$type=='Tumor']/sum(data$`Number of Cells`[data$type=='Tumor'])
 
  pC = ggplot(data=data, aes(x=Clusters, y=`Proportions of Cells`,fill=Annotation)) +
    geom_bar(position="stack", stat="identity") + scale_fill_brewer(palette="Set1") + facet_wrap(~type, scales = "free_x",nrow=2) + theme(plot.title = element_text(face = "bold")) + ggtitle('Cell annotation per cluster and tissue type') 
  pC$labels$fill = 'Annotations'
  
  pdf(file.path(params$datapath,paste0("../results/Figure",figure[i],"_UMAP_",methods[i],".pdf")),width = 11,height = 7)
  (pA+pB)/(pC) + plot_annotation(tag_levels = 'A')
  dev.off()
}
```

#DEGs general 
```{r degs}
degs = data.frame(cell_type = c('Stroma','Immune','Epithelial','Endothelial'), `Single Cells\nvs Single Nuclei` = rep(0,4),`Single Cells\nvs Depleted` = 0, `Single Nuclei\nvs Depleted` = 0,check.names = F)

adeno.combined.sct <- SetIdent(adeno.combined.sct, value = 'predicted.ann_level_1')

for(j in seq_along(degs$cell_type)) {
  deg1 <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "Single Cells",ident.2 = "Single Nuclei", group.by = 'method',assay = 'integrated',subset.ident = degs$cell_type[j],max.cells.per.ident=1000)

  deg2 <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "Single Cells",ident.2 = "Single Cells (depletion)", group.by = 'method',assay = 'integrated',subset.ident = degs$cell_type[j],max.cells.per.ident=1000)

  deg3 <- FindMarkers(adeno.combined.sct,slot = 'scale.data', ident.1 = "Single Cells (depletion)",ident.2 = "Single Nuclei", group.by = 'method',assay = 'integrated',subset.ident = degs$cell_type[j],max.cells.per.ident=1000)

degs[j,2] = nrow(deg1[deg1$p_val_adj<0.05,])
degs[j,3] = nrow(deg2[deg2$p_val_adj<0.05,])
degs[j,4] = nrow(deg3[deg3$p_val_adj<0.05,])

print(paste0(j,'---', Sys.time()))
}

degs = degs %>% pivot_longer(cols = 2:4,names_to = "Comparisons",values_to = "Number of DEGs")

p_degs = ggplot(data=degs, aes(x=Comparisons, y=`Number of DEGs`,fill=Comparisons)) +
    geom_bar(stat="identity") + 
    facet_wrap(~cell_type)

pdf(file.path(params$datapath,'../results/Figure9_degs.pdf'),width = 9,height = 6)
p_degs
dev.off()

```

#heatmap/dotplot ? 
```{r,eval = F}
markers = read.csv('scRNA/scRNA/data/marker_genes_April2023.txt',header = F,row.names = 1)

###epithelial
epi = subset(adeno.combined.sct,subset = predicted.ann_level_1= 'Epithelial')
DotPlot(epi, features = unlist(markers[c(3,6,12,15,36,43,49),]),
        split.by = "predicted.ann_finest_level") + RotatedAxis()
```


# session info  
```{r session, message= T}
###session
sessionInfo()
```

