---
title: "Single Nuclei vs Single cells (non-depleted) -- Normal tissue, 2 patients"
author: "Sébastien Renaut (sebastien.renaut@criucpq.ulaval.ca)"
date: "`r Sys.time()`"
output:
  html_document:
    number_sections: T
params:
  datapath: 'C:/Users/renseb01/Documents/scRNA/scRNA/data'
  outputpath: '../..' 
---

# Setup  
* This is a scRNA project 
* Note that you can create this document by running:  
  * `Rscript -e "rmarkdown::render('scRNA_seurat_lung_adenocarcino_vSCT_merge_intergrated.RMd',params = list())"` and `knit`-it directly in `RStudio`.
  
```{r, timehook,echo=F}
knitr::knit_hooks$set(timehook = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("\n\n\nTime for this code chunk to run:",signif(res,3),attr(res,'units'))
    }
  }
}))
```

```{r, setup,echo=FALSE}
knitr::opts_knit$set(root.dir = params$datapath)
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, cache=FALSE, cache.lazy=FALSE)
```

```{r, requirements}
library(Seurat) #scRNA analyses
library(sctransform) #data normalization
library(scater) #outlier values predictions
library(scAnnotatR) #cell type annotation (devtools::install_github('grisslab/scAnnotatR'))
library(DoubletFinder) #find doublets? 

library(vegan) #diversity indices

library(dplyr) #data handling
library(patchwork) #data viz
library(ggplot2) #data viz
library(DT) #data viz
library(plotly)
source('../R/doubletfinder.R')
source('../R/azimuth_analysis.R')
```

# Loading data  
  * `barcodes.tsv.gz`, `features.tsv.gz` and `matrix.mtx.gz` files come directly from  `cell ranger`.


```{r filenames,message  = T, fig.width = 11, fig.height=8,echo=T}  
filenames =  list.files('.',pattern = '-C[ST]T-',full.names = F)
filepaths = file.path(getwd(),filenames,'filtered_feature_bc_matrix')
filenames
```


```{r qc,message  = T, fig.width = 11, fig.height=8,echo=F, timehook = TRUE}  
#create seurat objects
seurat.objects = list()
l = length(filenames)

#create filters objects
QC_filters=data.frame(nCount_min=rep(0,l),nCount_max=rep(0,l),
                    nFeature_RNA_min=rep(0,l),nFeature_RNA_max=rep(0,l),
                    percent.mt_min=rep(0,l),percent.mt_max=rep(0,l))


for(i in 1:length(filenames)) {
#for(i in 1:1) {
  # Read 10x data
  cts <- Read10X(data.dir = filepaths[i])
  seurat.objects[[i]] = CreateSeuratObject(counts = cts)
  print(seurat.objects[[i]])
  seurat.objects[[i]][["percent.mt"]] <- PercentageFeatureSet(seurat.objects[[i]], pattern = "^MT-")

  # Filters
  QC_filters[i,1:2] = attr(isOutlier(seurat.objects[[i]]@meta.data$nCount),'thresholds')
  QC_filters[i,3:4] = attr(isOutlier(seurat.objects[[i]]@meta.data$nFeature_RNA),'thresholds')
  QC_filters[i,5:6] = attr(isOutlier(seurat.objects[[i]]@meta.data$percent.mt),'thresholds')
  
  # Subset
  seurat.objects[[i]] <- subset(seurat.objects[[i]], subset = percent.mt <  QC_filters[i,6])

  # Generate metadata
  temp = strsplit(filenames[i],split = '_')[[1]]
  if(temp[2]=='all')temp[2]='Single Cells'
  if(temp[2]=='depletion')temp[2]='Single Cells (depletion)'
  if(temp[2]=='SN')temp[2]='Single Nuclei'
  meta = c(filenames[i],temp[2],ifelse(regexpr('CST',temp[1])>0,'Normal','Tumor'),strsplit(temp[1],fixed = T,'-')[[1]][2])
  seurat.objects[[i]]@meta.data$samplename = temp[1]
  seurat.objects[[i]]@meta.data$method = temp[2]
  seurat.objects[[i]]@meta.data$type = ifelse(regexpr('CST',temp[1])>0,'Normal','Tumor')
  seurat.objects[[i]]@meta.data$patient = ifelse(regexpr('^4',temp[1])>0,'P4','P6')
  
  # Normalisation
  seurat.objects[[i]] <- SCTransform(seurat.objects[[i]], method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = F)
  
  print(paste0(i,' --- Time: ',Sys.time(),' -- seurat object ready (subsetting, normalisation)'))
}

names(seurat.objects) =  filenames

print(seurat.objects)
```


# Doublet Finder
  * find and remove doublet (assuming a 5% doublet rate).
```{r doublet finder, echo=T, message=FALSE, warning=FALSE,include = FALSE, timehook=T}
#precomputed pk's for doubletfinder
pK_empirical = c(0.030,0.010,0.020,0.300,0.300,0.001,0.210,0.060,0.010,0.090,0.240,0.040)

for(i in 1:length(filenames)) {
  #doublets finder
  doubletfinder.out = doubletfinder_wrapper(seurat.objects[[i]], pK_empirical = pK_empirical[i])
  seurat.objects[[i]] = doubletfinder.out[[1]]
  pK_empirical[i] =  doubletfinder.out[[2]]
}
print(seurat.objects)
```

```{r doublet finder results, message=T}
#precomputed pk's for doubletfinder
print(seurat.objects)
summary_table = data.frame(sample = filenames,
                           type = '' ,
                           method = '',
                           patient = '',
                           nb_cells = '',
                           nb_genes = '')

for(i in seq_along(filenames)){
  summary_table$type[i] = unique(seurat.objects[[i]]$type)
  summary_table$method[i] = unique(seurat.objects[[i]]$method)
  summary_table$patient[i] =  unique(seurat.objects[[i]]$patient)
  summary_table$nb_cells[i] =  ncol(seurat.objects[[i]])
   summary_table$nb_genes[i] =  nrow(seurat.objects[[i]])
}

summary_table = summary_table[order(summary_table$type,summary_table$method),]

write.table(summary_table,file.path(params$datapath,"../results/summary_table.csv"),row.names = F, quote = F,sep = '\t')

```


# Azimuth Annotation
  * Annotation based on [HLCA version 2](https://app.azimuth.hubmapconsortium.org/app/human-lung-v2) from Azimuth.
  * 58 lung specific cell types.
```{r azimuth annotation,include=FALSE}
  # Cell Type Identification
for(i in 1:length(filenames)) {
  seurat.objects[[i]] = azimuth_annotation(seurat.objects[[i]],annotation_threshold = 0.5)
}
```


# Filters defined with isOutlier()
```{r DTfilter,echo =F, message = T} 
rownames(QC_filters) = filenames
QC_filters = signif(QC_filters,4)
datatable(QC_filters,caption = htmltools::tags$caption(htmltools::strong("Filters as calculated by isOutlier()"), style="color:darkred"), options = list(dom = 't'))
```

# session info  
```{r session, message= T}
###session
saveRDS(seurat.objects, file = file.path(params$datapath,"adeno.12samples.rds"))
sessionInfo()
```






